Linear analysis of the vertical shear instability in astrophysical disks

### 2015-02-12 appendix.tex
spell-checked first rough draft
minor edits

\section{Applicability of the vertically global shearing box}\label{global_corr}
In \S\ref{analytical} we used the vertically global shearing box
\citep[VGSB,][]{mcnally14} to make analytical progress. % Here, we discuss limitations of this approach.   
The main difference between the linear problem in the VGSB 
and the radially-local approximation
(Eq. \ref{lin_mass}---\ref{lin_energy}) adopted in our numerical study
is the absence of  terms associated with the radial disk structure in
the former framework. These terms appear in the energy, mass and radial momentum
equations and are reproduced here for convenience,
% Using the notation and disk models in the main text, the
% energy, density and $r$ momentum equations become 
\begin{align}
  \ii \sigma W  &=c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
      z}\right|_0 \delta v_z + \frac{\gamma}{\Gamma} \frac{d\delta
    v_z}{dz}\right) + \left[\ii k_xc_s^2(r_0,z)
  \frac{\gamma}{\Gamma} + \hat{g}_c\left.\frac{1}{\rho}\frac{\p P}{\p
      r}\right|_0\right]\delta v_x  +
\frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
\ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_r + \hat{g}_c
  \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_x + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
      z}\right|_0 \delta v_z + \frac{d\delta
    v_z}{dz}\right),\label{gcorr_terms2} \\
\ii \sigma \delta v_x & = \ii k_x W  -
\hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
  r}\right|_0Q - 2\Omega(r_0,z)\delta v_y,\label{gcorr_terms3}
\end{align}
where subscript $0$ denotes evaluation at $r=r_0$.

@@

\subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
An important caveat of the VGSB is that it cannot be used in the
adiabatic limit. We explain this by setting  $\beta\to\infty$ in
Eq. \ref{vertiso_gov_nondim}, which can then be written as 
% For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
% simply use the Soilberg-Hoiland critera to assess stability
% (\S\ref{solberg}). Nevertheless, we briefly consider this case in the
% local framework to demonstrate its limitations. For $\beta\to\infty$
% the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
% be written as   
\begin{align}
  0 =\dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime
  +\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
    -\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
\end{align}
We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
integrate vertically, assume boundary terms vanish when integrating by
parts, to obtain
\begin{align}
  \hat{\sigma}^2\left(\frac{1}{\gamma} +
    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
  v_z|^2 d\zhat 
  =&  \left(\frac{\gamma-1}{\gamma}\right)
  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
  v_z)^\prime|^2 d\zhat\notag\\
&+
  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
    2}|\delta v_z|^2 d\zhat
+ \ii\epsilon q \hat{k}
  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
  \delta v_z d\zhat.\label{adia_integral}
\end{align}
When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
expected, a sub-adiabatically stratified disk is stable in the absense
of vertical shear.

@@

\section{Axisymmetric instability in vertically 
  isothermal disks in global cylindrical geometry}\label{global_vertiso}
Consider a locally isothermal disk in 3D such that $P=c_s^2\rho$ at
all times, so in the perturbed state $\delta P =
c_s^2\delta\rho$, with $c_s$ begin a prescribed function of position. 
The linearized fluid equations in cylindrical
geometry is then 
\begin{align}
 \ii\sigma \delta v_r + 2\Omega\delta v_\phi &= c_s^2\frac{\p}{\p r}\left(\frac{\delta\rho}{\rho}\right),\\
 \ii\sigma \delta v_\phi &= \frac{1}{r}\delta\bm{v}\cdot\nabla j,\\
 \ii\sigma \delta v_z &= c_s^2\frac{\p}{\p
   z}\left(\frac{\delta\rho}{\rho}\right),\\
 \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right).
\end{align}
We can eliminate $\delta v_\phi$ from these equations to obtain
\begin{align}
  -\sigma^2\rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) +
  \kappa^2\rho|\delta v_r|^2 + \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
  \delta v_r^*\delta v_z =
  c_s^2\rho\delta\bm{v}^*\cdot\nabla\left(\frac{\ii\sigma\delta\rho}{\rho}\right)
  =c_s^2\rho\delta\bm{v}^*\cdot\nabla\left[\frac{1}{\rho}\nabla\cdot\left(\rho\delta\bm{v}\right)\right].\label{global_integral} 
\end{align}
At this point if we make the anelastic approximation by assuming
$\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this
equation \citep[as argued by][]{nelson13}, 
then the only contribution to instability comes from the third term on
the left hand side, i.e. the explicit vertical shear term
$\p_zj$. This is the same conclusion reached in
\S\ref{integral_relation}.

@@

If we integrate Eq. \ref{global_integral} over the fluid volume 
and assume boundary terms vanish when integrating by parts, we obtain 
\begin{align}
  -\sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV +
  \int \kappa^2\rho|\delta v_r|^2 dV + \int \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
  \delta v_r^*\delta v_z dV = -\int
  \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
  dV - \int \nabla\cdot\left(\rho\delta\bm{v}\right) \delta
  \bm{v}^*\cdot\nabla c_s^2.
\end{align}
Note that the treatment so far is in fact valid for any prescribed
$c_s(r,z)$. However, to proceed further we specialize to the case
$c_s=c_s(r)$ for simplicity, i.e. vertically isothermal
disks. Recalling $j=r^2\Omega$ and $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
c_s^2$  for a vertically isothermal disk, we obtain 
\begin{align}
  \sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV =
  \int \kappa^2\rho|\delta v_r|^2 dV + \int
  \frac{c_s^2(r)}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
  dV + \int \left[ \rho \left(\nabla\cdot\delta\bm{v}\right)\delta
    v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2(r)}{dr}  dV.
\end{align}
We see that instability is only possible if
$dc_s^2/dr\neq0$. Furthermore, since density and temperature typically
decrease outwards in astrophysical disks, instability is associated
with the term proportional to $\nabla\cdot\delta\bm{v}$. It is also
clear that $\sigma^2$ will be complex in general.

@@

\section{Fiducial model for a protoplanetary disk}\label{mmsn}
For results application in \S\ref{application} we use the disk model
described in \cite{chiang10}. This disk model orbits a Solar-mass star and 
has the surface density distribution
\begin{align}\label{mmsn_sigma}
  \Sigma = 2200
  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-3/2}\mathrm{g}\,\mathrm{cm}^{-2},  
\end{align}
and the temperature profile
\begin{align}\label{mmsn_temp}
  T = 120\hat{T}\left(\frac{r}{\mathrm{AU}}\right)^{-3/7} \mathrm{K},
\end{align}
which implies $q=-3/7$. In the above expressions, $\hat{\Sigma}$ and
$\hat{T}$ are dimensionless coefficients used to scale the model
relative to the MMSN. Their nomial values are unity. By assuming a
vertically isothermal disk, we deduce the disk aspect-ratio 
\begin{align}\label{mmsn_epsilon}
  \epsilon =
  3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}, 
\end{align}
and the mid-plane density distribution 
\begin{align}
%  \rho_0 = 2.7\times10^{-9}
%  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
\rho_0 = 1.7\times10^{-9}
  \hat{\Sigma}\left(\frac{\hat{T}}{\mu}\right)^{-1/2}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},
\end{align}
which implies $p=-39/14$. 
%0.033576258
In addition, we
use the opacity model {\bf(reference?)}
 \begin{align}
   \kappa_d &= 2 \hat{\kappa}_d \left(\frac{T}{100\mathrm{K}}\right)^2
   \mathrm{cm}^2\,\mathrm{g}
    =
   2.88\hat{\kappa}_d\hat{T}^2\left(\frac{r}{\mathrm{AU}}\right)^{-6/7}\mathrm{cm}^2\,\mathrm{g}^{-1},   
 \end{align}
where the second equality follows from the model temperature profile
above, and $\hat{\kappa}_d$ has similar meaning as $\hat{\Sigma}$ and
$\hat{T}$.

>>>>

\section{Applicability of the vertically global shearing box}\label{global_corr}
In \S\ref{analytical} we used the vertically global shearing box
\citep[VGSB,][]{mcnally14} to make analytic progress. % Here, we discuss limitations of this approach.   
The main difference between the linear problem in the VGSB 
and the radially-local approximation
(Eq. \ref{lin_mass}---\ref{lin_energy}) adopted in our numerical study
is the absence of  terms associated with the radial disk structure in
the former framework. These terms appear in the energy, mass and radial momentum
equations and are reproduced here for convenience,
% Using the notation and disk models in the main text, the
% energy, density and $r$ momentum equations become 
\begin{align}
  \ii \sigma W  &=c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
      z}\right|_0 \delta v_z + \frac{\gamma}{\Gamma} \frac{d\delta
    v_z}{dz}\right) + \left[\ii k_xc_s^2(r_0,z)
  \frac{\gamma}{\Gamma} + \hat{g}_c\left.\frac{1}{\rho}\frac{\p P}{\p
      r}\right|_0\right]\delta v_x  +
\frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
\ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_r + \hat{g}_c
  \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_x + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
      z}\right|_0 \delta v_z + \frac{d\delta
    v_z}{dz}\right),\label{gcorr_terms2} \\
\ii \sigma \delta v_x & = \ii k_x W  -
\hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
  r}\right|_0Q - 2\Omega(r_0,z)\delta v_y,\label{gcorr_terms3}
\end{align}
where subscript $0$ denotes evaluation at $r=r_0$.

@@

\subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
An important caveat of the VGSB is that it cannot be used in the
adiabatic limit. We explain this by setting  $\beta\to\infty$ in
Eq. \ref{vertiso_gov_nondim}, which can then be written as 
% For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
% simply use the Soilberg-Hoiland critera to assess stability
% (\S\ref{solberg}). Nevertheless, we briefly consider this case in the
% local framework to demonstrate its limitations. For $\beta\to\infty$
% the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
% be written as   
\begin{align}
  0 =\dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime
  +\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
    -\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
\end{align}
We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
integrate vertically, assume boundary terms vanish when integrating by
parts, to obtain
\begin{align}
  \hat{\sigma}^2\left(\frac{1}{\gamma} +
    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
  v_z|^2 d\zhat 
  =&  \left(\frac{\gamma-1}{\gamma}\right)
  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
  v_z)^\prime|^2 d\zhat\notag\\
&+
  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
    2}|\delta v_z|^2 d\zhat
+ \ii\epsilon q \hat{k}
  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
  \delta v_z d\zhat.\label{adia_integral}
\end{align}
When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
expected, a sub-adiabatically stratified disk is stable in the absence
of vertical shear.

@@

\section{Axisymmetric instability in vertically 
  isothermal disks in global cylindrical geometry}\label{global_vertiso}
Consider a locally isothermal disk in 3D such that $P=c_s^2\rho$ at
all times, so in the perturbed state $\delta P =
c_s^2\delta\rho$, with $c_s$ begin a prescribed function of position. 
The axisymmetric linearized fluid equations in cylindrical
geometry is then 
\begin{align}
 \ii\sigma \delta v_r + 2\Omega\delta v_\phi &= c_s^2\frac{\p}{\p r}\left(\frac{\delta\rho}{\rho}\right),\\
 \ii\sigma \delta v_\phi &= \frac{1}{r}\delta\bm{v}\cdot\nabla j,\\
 \ii\sigma \delta v_z &= c_s^2\frac{\p}{\p
   z}\left(\frac{\delta\rho}{\rho}\right),\\
 \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right),
\end{align}
and we recall $j=r^2\Omega$. We can eliminate $\delta v_\phi$ from these equations to obtain
\begin{align}
  -\sigma^2\rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) +
  \kappa^2\rho|\delta v_r|^2 + \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
  \delta v_r^*\delta v_z =
  c_s^2\rho\delta\bm{v}^*\cdot\nabla\left(\frac{\ii\sigma\delta\rho}{\rho}\right)
  =c_s^2\rho\delta\bm{v}^*\cdot\nabla\left[\frac{1}{\rho}\nabla\cdot\left(\rho\delta\bm{v}\right)\right].\label{global_integral} 
\end{align}
At this point if we make the anelastic approximation by assuming
$\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this
equation \citep[as argued by][]{nelson13}, 
then the only contribution to instability comes from the third term on
the left hand side, i.e. the explicit vertical shear term
$\p_zj$. This is the same conclusion reached in
\S\ref{integral_relation}.

@@

If we integrate Eq. \ref{global_integral} over the fluid volume 
and assume boundary terms vanish when integrating by parts, we obtain 
\begin{align}
  -\sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV +
  \int \kappa^2\rho|\delta v_r|^2 dV + \int \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
  \delta v_r^*\delta v_z dV = -\int
  \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
  dV - \int \nabla\cdot\left(\rho\delta\bm{v}\right) \delta
  \bm{v}^*\cdot\nabla c_s^2.
\end{align}
Note that the treatment so far is in fact valid for any prescribed
$c_s(r,z)$. However, to proceed further we specialize to the case
$c_s=c_s(r)$ for simplicity, i.e. vertically isothermal
disks. Recalling $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
c_s^2$  for a vertically isothermal disk (Eq. \ref{vertical_shear}), we obtain 
\begin{align}
  \sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV =
  \int \kappa^2\rho|\delta v_r|^2 dV + \int
  \frac{c_s^2(r)}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
  dV + \int \left[ \rho \left(\nabla\cdot\delta\bm{v}\right)\delta
    v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2(r)}{dr}  dV.
\end{align}
We see that instability is only possible if
$dc_s^2/dr\neq0$. Furthermore, since density and temperature typically
decrease outwards in astrophysical disks, instability is associated
with the term proportional to $\nabla\cdot\delta\bm{v}$. It is also
clear that $\sigma^2$ will be complex in general.

@@

\section{Fiducial model for a protoplanetary disk}\label{mmsn}
For results application in \S\ref{application} we use the disk model
described in \cite{chiang10}. This disk model orbits a Solar-mass star and 
has the surface density distribution
\begin{align}\label{mmsn_sigma}
  \Sigma = 2200
  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-3/2}\mathrm{g}\,\mathrm{cm}^{-2},  
\end{align}
and the temperature profile
\begin{align}\label{mmsn_temp}
  T = 120\hat{T}\left(\frac{r}{\mathrm{AU}}\right)^{-3/7} \mathrm{K},
\end{align}
which implies $q=-3/7$. In the above expressions, $\hat{\Sigma}$ and
$\hat{T}$ are dimensionless coefficients used to scale the model
relative to the MMSN. Their nominal values are unity. By assuming a
vertically isothermal disk, we deduce the disk aspect-ratio 
\begin{align}\label{mmsn_epsilon}
  \epsilon =
  3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}, 
\end{align}
and the mid-plane density distribution 
\begin{align}
%  \rho_0 = 2.7\times10^{-9}
%  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
\rho_0 = 1.7\times10^{-9}
  \hat{\Sigma}\left(\frac{\hat{T}}{\mu}\right)^{-1/2}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},
\end{align}
which implies $p=-39/14$. 
%0.033576258
In addition, we
use the opacity model {\bf(reference?)}
 \begin{align}
   \kappa_d &= 2 \hat{\kappa}_d \left(\frac{T}{100\mathrm{K}}\right)^2
   \mathrm{cm}^2\,\mathrm{g}
    =
   2.88\hat{\kappa}_d\hat{T}^2\left(\frac{r}{\mathrm{AU}}\right)^{-6/7}\mathrm{cm}^2\,\mathrm{g}^{-1},   
 \end{align}
where the second equality follows from the model temperature profile
above, and $\hat{\kappa}_d$ has similar meaning as $\hat{\Sigma}$ and
$\hat{T}$.

### 2015-02-12 paper.tex
minor edits

%what we're doing
In this work we study the linear VSI in radially-local,
vertically-global models of astrophysical disks. We include an energy
equation with a source term that allows the thermodynamic response of
perturbations between varied from isothermal to adiabatic. Our
formulation generalizes previous linear calculations of the 
VSI carried out by \cite{nelson13} and \cite{mcnally14} which assume
isothermal perturbations. Our models are also an extension of early
analysis of \cite{urpin98} and \cite{urpin03} who consider
vertically-localized disturbances. Our vertically-global approach is
necessary to study disturbances with vertical wavelengths
comparable to or larger than the disk scale height. We thus focus on
vertically-global perturbations, which were found by \cite{nelson13}
to dominate their numerical simulations.

>>>>

%what we're doing
In this work we study the linear VSI in radially-local,
vertically-global models of astrophysical disks. We include an energy
equation with a source term that allows the thermodynamic response of
perturbations to be varied smoothly between isothermal and adiabatic. Our
formulation generalizes previous linear calculations of the 
VSI carried out by \cite{nelson13} and \cite{mcnally14} which assume
isothermal perturbations. Our models are also an extension of early
analysis of \cite{urpin98} and \cite{urpin03} who consider
vertically-localized disturbances. Our vertically-global approach is
necessary to study disturbances with vertical wavelengths
comparable to or larger than the disk scale height. Such disturbances were 
found by \cite{nelson13} to dominate their numerical simulations. 
We thus focus on vertically-global perturbations.

### 2015-02-12 summary.tex
minor edits

In accordance with the above studies, which show that the VSI applies
to perturbations with small radial lengthscales, we applied the 
radially-local approximation in the linearized fluid equations. For
analytic discussion only, we further simplified the linear problem by
ignoring the background radial disk structure. This is equivalent to adopting
the `vertically-global shearing box' (VGSB) formalism developed by
\cite{mcnally14}, who showed the linear VSI studied by
\cite{nelson13} in a locally isothermal disk can be captured by the
VGSB. We find the VGSB framework is appropriate for $\beta\lesssim1$. 
This is convenient because previous works show that rapid thermal
relaxation ($\beta\ll1$) is generally required for the VSI. However,
the VGSB framework is not valid in the adiabatic limit.

@@

We also solved the linear stability problem numerically using a
pseudo-spectral approach. For numerical calculations we accounted for
the background radial disk structure for consistency. We focused
on the fundamental mode to compare with our analytical
discussion, and because \cite{nelson13} found that this mode 
eventually dominate their simulations. Our numerical solutions are
mostly consistent with our analytic expectations, and we verified the
dependence of $\beta_\mathrm{crit}$ on disk parameters.

>>>>

In accordance with the above studies, which show that the VSI occurs for 
perturbations with small radial lengthscales, we applied the 
radially-local approximation in the linearized fluid equations. For
analytic discussion only, we further simplified the linear problem by
ignoring the background radial disk structure. This is equivalent to adopting
the `vertically-global shearing box' (VGSB) formalism developed by
\cite{mcnally14}, who showed the linear VSI studied by
\cite{nelson13} in a locally isothermal disk can be captured by the
VGSB. We find the VGSB framework is appropriate for $\beta\lesssim1$. 
This is convenient because previous works show that rapid thermal
relaxation ($\beta\ll1$) is generally required for the VSI. However,
the VGSB framework is not valid in the adiabatic limit.

@@

We also solved the linear stability problem numerically using a
pseudo-spectral approach. For numerical calculations we accounted for
the background radial disk structure for consistency. We focused
on the fundamental mode to compare with our analytical
discussion, and because \cite{nelson13} found that this mode 
eventually dominate their simulations. Our numerical solutions are
mostly consistent with our analytic expectations, including the 
dependence of $\beta_\mathrm{crit}$ on disk parameters.
